---
title: "Running gene-set anaysis with `piano`"
author:
- name: "Leif Varemo Wigge"
  affiliation: 
  - Department of Biology and Biological Engineering, Chalmers University of Technology
  - National Bioinformatics Infrastructure Sweden (NBIS), SciLifeLab
package: "piano"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Running gene-set anaysis with piano}
  %\VignetteDepends{piano}
  %\VignetteKeywords{piano}
  %\VignettePackage{piano}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Quick start

1. Load piano and example data:

```{r}
library(piano)
data("gsa_input")
```

2. Take a look at the structure of the input data:

```{r}
head(gsa_input$gsc,10)
head(gsa_input$pvals, 10)
head(gsa_input$directions, 10)
```

3. Load gene-set collection and take a look at the resulting object:

```{r}
geneSets <- loadGSC(gsa_input$gsc)
geneSets

```

4. Run gene-set analysis:

```{r results='hide', message=F}
gsares <- runGSA(gsa_input$pvals,
                 gsa_input$directions,
                 gsc = geneSets,
                 nPerm = 500) # set to 500 for fast run
```
*Note: `nPerm` was set to 500 to get a short runtime for this vignette, in reality use a higher number, e.g. 10,000 (default).*

5. Explore the results in an interactive Shiny app:
```{r eval=F}
exploreGSAres(gsares)
```

This opens a browser window with an interactive interface where the results can be explored in detail.

# Session info
Here is the output of `sessionInfo()` on the system on which this document was compiled.

```{r echo=FALSE}
sessionInfo()
```


